Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-25 cav2017/cav2017.tex
wordings

We describe and evaluate two approaches for synthesizing digital controllers
for state-space physical plants, both based on the CounterExample Guided
Inductive Synthesis (CEGIS)~\cite{sketch} architecture, recently of wide use
for program synthesis.  We provide two alternative automated procedures to
synthesize safe digital state-feedback controls and prove their soundness by
encompassing errors due to digitalization and quantization effects that are
present in the digital controller interacting with the continuous plant. 
Both models synthesize sound safe controllers given a discrete-time model of
the closed-loop system.

{\em The first approach} uses a {\em na\"ive} technique that starts by
devising a digital controller that stabilizes the system whilst remaining
safe for a pre-selected time horizon and a single initial state; then, it
verifies its potential safety by unfolding the dynamics of the system,
considering the full hyper-cube of initial states, and checking for a {\em
completeness threshold}, i.e., the number of iterations required to
sufficiently unwind the closed-loop state-space system such that the
boundaries are not violated for any larger number of iterations.  As it
requires explicit unfolding up to the completeness threshold, this approach
may suffer from state-space explosion.

Instead of explicitly unfolding the dynamics, {\em the second approach}
employs {\em abstract acceleration}~\cite{cattaruzza2015unbounded} to
implicitly evaluate all possible progressions of the system simultaneously
and thus avoid the state-space explosion problem.  Additionally, the second
approach uses {\em abstraction refinement} enabling us to always start with
a very simple description regardless of the complexity of the overall
dynamics, and only expand to more complex models when a solution cannot be
found.

@@

During any given ADC conversion, the continuous signal will be sampled in the real domain and transformed by $\mathcal{F}\langle I_{c},F_{c} \rangle (x)$ (assuming the ADC discretization is the same as the digital implementation). this sampling uses a threshold which is defined by the less significant bit ($q_{c}=2^{-F_c}$) of the ADC and some non-linearities of the circuitry. The overall conversion is
$$\mathcal{F}\langle I_{c},F_{c} \rangle(y(t)) = y_k : y_k \in \left[y(t)-\frac{q_{c}}{2}\ \ \ \ y(t)+\frac{q_{c}}{2}\right]$$
If we denote the error in the conversion as $\nu_k=y_k-y(t) : t=nk$ then we may define some bounds for it $\nu_k \in [-\frac{q_{c}}{2}\ \ \frac{q_{c}}{2}]$.

We will assume, for the purposes of this analysis, that the domain of the ADC is that of the digital controller (\emph{ie} the quantizer includes any digital gain added in the code).
The process of quantization in the DAC is similar except that it is calculating $\mathcal{F}\langle I_{dac},F_{dac} \rangle (\mathcal{F}\langle I_{c},F_{c} \rangle (x)) $. If these domains are the same ($I_{c}=I_{dac},F_{c}=F_{dac}$), or if the DAC resolution in higher than the ADCs, then the DAC quantization error is equal to zero.
From the above equations we can now define the ADC and DAC quantization noises, ${\nu_1}_k \in [-\frac{q_1}{2}\ \ \frac{q_1}{2}], {\nu_2}_k \in [-\frac{q_2}{2}\ \ \frac{q_2}{2}] : q_1=q_{c} \wedge q_2=q_{dac}$. Where $Q_1$ is the quantizer of the ADC and $Q_2$ the quantizer for the DAC (see fig. \ref{ssec:statefeedbackcontrol}).
Note that these bounds hold irrespective of whether the noise is correlated, hence we may use them to over-approximate the effect of the noise on the state space progression over time.
The resulting dynamics are:
\begin{align}
\label{eq:pre_quantization}
\vec{x}_{k+1} &= \vec{A}_d\vec{x}_k+\vec{B}_d(\vec{u}_k+{\vec{\nu}_2}_k),\\
u_k&=-K\vec{x}_{k}+{\vec{\nu}_1}_k. 
\end{align}

>>>>

We give two alternative approaches for synthesizing digital controllers for
state-space physical plants, both based on CounterExample Guided Inductive
Synthesis (CEGIS)~\cite{sketch}.  We prove their soundness by quantifying
errors caused digitalization and quantization effects that arise when
time/value-discrete digital controller interacts with the continuous plant.

{\em The first approach} uses a {\em na\"ive} technique that starts by
devising a digital controller that stabilizes the system while remaining
safe for a pre-selected time horizon and a single initial state; then, it
verifies unbounded safety by unfolding the dynamics of the system,
considering the full hyper-cube of initial states, and checking a {\em
completeness threshold}, i.e., the number of iterations required to
sufficiently unwind the closed-loop state-space system such that the
boundaries are not violated for any larger number of iterations.  As~it
requires explicit unfolding up to the completeness threshold, this approach
is computationally expensive.

Instead of explicitly unfolding the dynamics, {\em the second approach}
employs {\em abstract acceleration}~\cite{cattaruzza2015unbounded} to
implicitly evaluate all possible progressions of the system simultaneously. 
Additionally, the second approach uses {\em abstraction refinement},
enabling us to always start with a very simple description regardless of the
complexity of the overall dynamics, and only expand to more complex models
when a solution cannot be found.

@@

During any given ADC conversion, the continuous signal will be sampled in the real domain and transformed by $\mathcal{F}\langle I_{c},F_{c} \rangle (x)$. this sampling uses a threshold which is defined by the less significant bit ($q_{adc}=2^{-F_{adc}}$) of the ADC and some non-linearities of the circuitry. The overall conversion is
$$\mathcal{F}\langle I_{c},F_{c} \rangle(y(t)) = y_k : y_k \in \left[y(t)-\frac{q_{c}}{2}\ \ \ \ y(t)+\frac{q_{adc}}{2}\right]$$
If we denote the error in the conversion as $\nu_k=y_k-y(t) : t=nk$ then we may define some bounds for it $\nu_k \in [-\frac{q_{c}}{2}\ \ \frac{q_{c}}{2}]$.

We will assume, for the purposes of this analysis, that the domain of the ADC is that of the digital controller (\emph{ie} the quantizer includes any digital gain added in the code).
The process of quantization in the DAC is similar except that it is calculating $\mathcal{F}\langle I_{dac},F_{dac} \rangle (\mathcal{F}\langle I_{c},F_{c} \rangle (x)) $. If these domains are the same ($I_{c}=I_{dac},F_{c}=F_{dac}$), or if the DAC resolution in higher than the ADCs, then the DAC quantization error is equal to zero.
From the above equations we can now infer that $\nu_1 \in [-\frac{q_1}{2}\ \ \frac{q_1}{2}], \nu_2 \in [-\frac{q_2}{2}\ \ \frac{q_2}{2}] : q_1=q_{c} \wedge q_2=q_{dac}$.
Note that these bounds hold irrespective of whether the noise is correlated, hence we may use them to over-approximate the effect of the noise on the state space progression over time.
The resulting dynamics are:
\begin{align}
\label{eq:pre_quantization}
\vec{x}_{k+1} &= \vec{A}_d\vec{x}_k+\vec{B}_d(\vec{u}_k+{\vec{\nu}_2}_k),\\
u_k&=-K\vec{x}_{k}+{\vec{\nu}_1}_k. 
\end{align}
